---
output: github_document
title: "A Bayesian Framework for Baseflow Estimation in Rivers"
author: "Edgar Santos-Fernandez, PhD"
bibliography: ref.bib
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r echo=F, warning=F, message=F}
.libPaths('C:\\1\\R6')
library(tidyverse)
library(RColorBrewer)
library(rstan)
library(bayesplot)
library(depmixS4)
library(plotly)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

```

```bayesflow``` estimate baseflow use Bayesian inference in R and Stan using different methods, including the Lyne and Hollick (@lyne1979stochastic, @ladson2013standard) and Eckhardt's filters (@eckhardt2008comparison). 


This approach allows:

* incorporating covariates e.g. precipitation 

* treat the filter parameters probabilistically and obtain probabilistic estimates of the baseflow,

* capture and propagate uncertainty, and better understand the underlying processes. 

* imputing missing data and fills the gaps on the go.


# Methods of baseflow estimation

Baseflow is the portion of streamflow that is sustained by groundwater discharge, rather than by surface water runoff or precipitation. 

The Lyne and Hollick filter separates the streamflow into two components: a quick response component that is associated with surface water runoff and a slow response component that is associated with groundwater discharge. 
The method involves applying a low pass filter to the flow data to separate the two components and then estimating the baseflow by summing the slow response component over a given period of time.

The filter is generally defined as follows. Let $y_t$ be the water level at time $t$.
The quick flow response is obtained using:

$$
f_t = \max (0, 
\alpha f_{t-1} + \frac{1+\alpha}{2}(y_{t} - y_{t-1}))   
$$

The $\alpha$ parameter  determines the smoothing of the filter. The baseflow is obtained using:

$$
b_t = y_t - f_t
$$


@eckhardt2008comparison suggested an improved method that has been reported to perform well on different scenarios  

$$
b_t = \frac{(1-\omega)\alpha b_{t-1} + (1-\alpha)\omega y_t}{1-\alpha \omega}
$$

where $\omega$ is a baseflow index, usually fixed at 0.8, 0.5 or 0.25 depending on the characteristics of the stream.





## Installation

And the development version from [GitHub](https://github.com/) with:

``` r
devtools::install_github("EdgarSantos-Fernandez/bayesflow")
```




